Enhanced Residual Context-based Networks for Image Outpainting

### 2020-05-13 tex/paper.tex
Added figs and eqs

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure} 
\usepackage{amsmath}

@@

\begin{abstract} 
Describe {\bf briefly} the computational problem you tackled, the approach you explored, the analyses you performed, and the results you obtained.
Point out advantages of your approach compared to prior approaches (if any).
\end{abstract}

\section{Introduction}
{\bf Problem before solution}:
Begin by describing a (computational) {\bf problem} and why the problem is {\bf important}. If previous research has
explored plausible solutions to solve the problem, then point out how these approaches may be lacking. Then briefly {\bf motivate}
why a new approach (the approach you explore in this paper) is worth exploring. Don't go into too much detail yet how it works -- that will
be left for Section \ref{sec:proposed_method}).

\subsection{Research contributions}
Describe clearly and concisely what your paper {\bf uniquely} contributes to the research community, i.e., that has never been done before.

\section{Related Work}
Describe previous approaches (in somewhat greater detail than the introduction) to tackling the problem. Pick any citation style you prefer and use it
consistently (here's a random citation \cite{mitchell80}). Explain how previous work may be lacking in certain ways, and use your critical review
of prior work to further motivate your research.

\section{Proposed Method}
\label{sec:proposed_method}
Describe the computational approach -- based obviously on neural networks -- that you took to tackle the problem.

\begin{table}[h] 
  \centering  
    \begin{tabular}{c|c|c}  
    \hline
    \textbf{layer} & \textbf{output size} & \textbf{parameters}\\ 
    \hline \hline
      Conv & 16$\times$192$\times$192 & 5$\times$5, stride=1 \\
    \hline
      ResBlock & 128$\times$96$\times$96 & 3$\times$3, stride=2 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=2 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Trans-Conv & 128$\times$96$\times$96 & 4$\times$4, stride=1 \\
    \hline
      ReLU & 128$\times$96$\times$96 & None \\
    \hline
      Conv & 128$\times$96$\times$96 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 128$\times$96$\times$96 & None \\
    \hline
      Trans-Conv & 64$\times$192$\times$192 & None \\
    \hline
      ReLU & 64$\times$192$\times$192 & None \\
    \hline
      Conv & 32$\times$192$\times$192 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 32$\times$192$\times$192 & None \\
    \hline
      Conv & 3$\times$192$\times$192 & 3$\times$3, stride=2 \\
    \hline
      Sigmoid & 3$\times$192$\times$192 & None \\
    \hline
    \end{tabular}

  \caption{Architecture of generator, including parameters. Trans-Conv is transposed convolution.} 
  \label{tab:1}
  \label{table1} 
\end{table}

\begin{table}[h] 
  \centering  
    \begin{tabular}{c|c}  
    \hline
    \textbf{Layer} & \textbf{Parameters}\\ 
    \hline \hline
      Conv & 3$\times$3, stride=\(s_{in}\) \\
    \hline
      ReLU & None \\
    \hline
      Conv & 3$\times$3, stride=\(s_{in}\) \\
    \hline
      ReLU & None \\
    \hline
      Add & None \\
    \hline
    \end{tabular}

  \caption{Architecture of residual block used in generator, including parameters. The residual block applies input strides \(s_{in}\) as convolution layer parameters. Add is an element-wise addition of \(x_{in}\) and output tensor of previous layer.} 
  \label{tab:2}
  \label{table2} 
\end{table}

\begin{table}[h] 
  \centering  
    \begin{tabular}{c|c}  
    \hline
    \textbf{Layer} & \textbf{Parameters}\\ 
    \hline \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2, inplace=True \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2, inplace=True \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2, inplace=True \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2, inplace=True \\
    \hline
      Conv & 3$\times$3, stride=1 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2, inplace=True \\
    \hline
      Conv & 3$\times$3, stride=1 \\
    \hline

    \end{tabular}

  \caption{Architecture of global and local discriminators $D_G$ and $D_L$, respectively.} 
  \label{tab:3}
  \label{table3} 
\end{table}

@@

Training is done for 50 epochs, with a fixed learning rate of $\alpha=0.0003$ and two Adam optimizers with $\beta_1=0.5, \beta_2=0.999$, following the same hyperparameters proposed in Van Hoorick 2020 \cite{van_hoorick_image_2020} The loss functions are as follows:
\begin{align}
  L_{rec} &= ||x-G(x)||_1 \\
  L_{adv} &= ||D(G(x))-1||_2^2 \\
  L_G &= \lambda_{rec}L_{rec} + \lambda_{adv}L_{adv} \\
  L_D &= ||D(x)-1||_2^2 + ||D(G(x))-0||_2^2
\end{align}

@@

\begin{align}
  \lambda_{adv}(n) = \begin{cases}
    0.001, & \text{if }n\leq10 \\
    0.005, & \text{if }10<n\leq30 \\
    0.015, & \text{otherwise}
  \end{cases}
  \label{eq:loss_adv}
\end{align}

@@

>>>>

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subcaption}
\usepackage{amsmath}

@@

\begin{abstract} 
Although humans perform well at predicting what exists beyond the boundaries of an image, deep models struggle to understand context and extrapolation through retained information. This task is known as image outpainting and involves generating realistic expansions of an image’s boundaries. Current models use generative adversarial networks to generate results which lack localized image feature consistency and appear fake. We propose two methods to improve this issue: the use of a local and global discriminator, and the addition of residual blocks within the encoding section of the network. Comparisons of our model and the baseline’s L1 loss, MSE loss, and qualitative differences reveal our model is able to naturally extend object boundaries and produce more internally consistent images compared to current methods but produces lower fidelity images.
\end{abstract}

\section{Introduction}
The use of generative adversarial models has led to many new developments in the computer vision field. The task known as image inpainting, which aims to enhance or restore quality to parts of an image, requires a model to retain the context of the image while completing missing regions of the image, and iteratively improve the process of the generations. The problem explored in this paper is a significantly more ambiguous task than its inpainting alternative, and is commonly known as image outpainting. For this task, the input image boundaries are extrapolated outside of the original content of the input based upon the context of the image.\cite{sabini_painting_2018}

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
    \centering
    \begin{subfigure}[b]{0.15\textwidth}
        \includegraphics[width=\textwidth]{figs/fig1/masked_input}
        \caption{Masked input}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
        \includegraphics[width=\textwidth]{figs/fig1/baseline}
        \caption{Output}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
        \includegraphics[width=\textwidth]{figs/fig1/ground_truth}
        \caption{GT}
    \end{subfigure}
  	\caption{Example of image outpainting.}
  	\label{fig:fig1}
\end{figure}

Image outpainting, sometimes referred to as image context interpretation and extrapolation, requires the context of the image to be retained by the encoder to accurately re-generate the known input image, then create new portions to append to the edges of the image as shown in Figure \ref{fig:fig1}. There exists research on this task using GANs, however one of the biggest issues with image outpainting is the new additions to the image, known as hallucinations, typically appear lower quality than the base image and therefore can easily be distinguished as fake by the human eye.\cite{sabini_painting_2018} Therefore, to effectively create higher quality hallucinations, better image evaluation methods are required to match the discrimination abilities of human interpretation.

\subsection{Research contributions}
We provide insight into the use of a local discriminator alongside a global discriminator to improve the quality of hallucinations, additionally evaluating the performance of residual blocks applied to the context encoder portion of our generative network.

\section{Related Work}
In this section, we briefly review the previous works relating to this paper in three sub-fields: Generative Adversarial Networks, Image Inpainting, and Image Outpainting.

\subsection{Generative Adversarial Networks}
Deep generative models have shown success in various tasks of image and video generation problems. By training the generator and discriminator in tandem, the generator can capture the real data distribution and create more realistic images given the latent input. However, GANs are often prone to collapse and are difficult to train. Therefore, some newer techniques have been applied attempting to reduce this issue such as the addition of residual blocks to the encoder component of the generator specifically researched within a super-resolution task context.\cite{lim_enhanced_2017}

\subsection{Image Inpainting}
Image inpainting is the task of recovering missing regions or enhancing regions within images. Pathak et al. demonstrate that context encoder can achieve realistic results for generation of novel images sections for inpainting purposes.\cite{pathak_context_2016} Recent improvements include applying global and local discriminators to improve the generated feature consistency within the missing sections of the input image by identifying poor generations only within the area of the inpainted sections.\cite{iizuka_globally_2017} Liu et al. more recently apply partial convolutions to improve overall consistency of generated images.\cite{liu_image_2018}

\section{Proposed Method}
In this section, we provide an overview of our architecture. Since the task is generative we use a GAN which other current methods employ. Our GAN is based on the architecture provided by Van Hoorick with two major additions: the first being the use of residual blocks in the generator, and the second being the use of two discriminators.\citep{van_hoorick_image_2020} Below we explore the exact architecture and some design decisions behind it.

\subsection{Generator}
We propose a GAN architecture that aims to encode the context of the image and use deconvolutional layers to then generate the resulting image while retaining feature consistency. We apply a residual model to the generative network, first convolving the image down into its feature space, then using deconvolution to upsample the image into its expanded version. Although typical models that contain residual blocks often include batch normalization within the blocks, we do not include this activation layer per demonstrations of improvement in the deblurring work of Nah et al.\cite{nah_deep_2018} The encoder section of the generative network contains multiple residual blocks in an attempt to improve image quality. Table \ref{tab:1} outlines the layers of our generative network, including both the encoder and decoder components.

\begin{table}[h] 
  \centering  
    \begin{tabular}{lcc}  
    \hline
    \textbf{layer} & \textbf{output size} & \textbf{parameters}\\ 
    \hline \hline
      Conv & 16$\times$192$\times$192 & 5$\times$5, stride=1 \\
    \hline
      ResBlock & 128$\times$96$\times$96 & 3$\times$3, stride=2 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=2 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ResBlock & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Conv & 256$\times$48$\times$48 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 256$\times$48$\times$48 & None \\
    \hline
      Trans-Conv & 128$\times$96$\times$96 & 4$\times$4, stride=1 \\
    \hline
      ReLU & 128$\times$96$\times$96 & None \\
    \hline
      Conv & 128$\times$96$\times$96 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 128$\times$96$\times$96 & None \\
    \hline
      Trans-Conv & 64$\times$192$\times$192 & None \\
    \hline
      ReLU & 64$\times$192$\times$192 & None \\
    \hline
      Conv & 32$\times$192$\times$192 & 3$\times$3, stride=1 \\
    \hline
      ReLU & 32$\times$192$\times$192 & None \\
    \hline
      Conv & 3$\times$192$\times$192 & 3$\times$3, stride=2 \\
    \hline
      Sigmoid & 3$\times$192$\times$192 & None \\
    \hline
    \end{tabular}

  \caption{Architecture of generator $G$, including parameters. Trans-Conv is transposed convolution.} 
  \label{tab:1} 
\end{table}

\begin{table}[h] 
  \centering  
    \begin{tabular}{lc}  
    \hline
    \textbf{Layer} & \textbf{Parameters}\\ 
    \hline \hline
      Conv & 3$\times$3, stride=$s_{in}$ \\
    \hline
      ReLU & None \\
    \hline
      Conv & 3$\times$3, stride=$s_{in}$ \\
    \hline
      ReLU & None \\
    \hline
      Add & None \\
    \hline
    \end{tabular}

  \caption{Architecture of residual block used in generator $G$, including parameters. The residual block applies input strides $s_{in}$ to the convolution layer. Add is an element-wise addition of block input tensor $x$ and output tensor of previous layer.} 
  \label{tab:2}
\end{table}

\subsection{Discriminator}
Our discriminator network $D$ consists of dual discriminators. One discriminator $D_G$ takes as input the entire generated image and applies a convolutional network to identify the image as real or generated, while the second discriminator $D_L$ is local, and is restricted solely to the boundaries between the input image and the hallucinations. These discriminator outputs are averaged, meaning our total discriminator network output is

\begin{align}
	D(x) &= \frac{D_L(\mu \times x) + D_G(x)}{2}
\end{align}

where $\mu$ is the binary mask which is originally applied to the training images.

The purpose of the local discriminator is to attempt to remove the clearly fake hallucinations that result from pixel noise or manifestations, that to the human eye are clearly fake. Since these happen in the boundary between the input image and the hallucinations, the local discriminator focuses on these areas alone as shown in Table \ref{tab:3}.

\begin{table}[h] 
  \centering  
    \begin{tabular}{lc}  
    \hline
    \textbf{Layer} & \textbf{Parameters}\\ 
    \hline \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2 \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2 \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2 \\
    \hline
      Conv & 3$\times$3, stride=2 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2 \\
    \hline
      Conv & 3$\times$3, stride=1 \\
    \hline
      InstanceNorm & None \\
    \hline
      LeakyReLU & negative slope=0.2 \\
    \hline
      Conv & 3$\times$3, stride=1 \\
    \hline
    \end{tabular}

  \caption{Architecture of global and local discriminators $D_G$ and $D_L$, respectively.} 
  \label{tab:3}
\end{table}

@@

Training is done for 50 epochs, with a fixed learning rate of $\alpha=0.0003$ and two Adam optimizers with $\beta_1=0.5, \beta_2=0.999$, following the same hyperparameters proposed in Van Hoorick 2020 \cite{van_hoorick_image_2020} The loss functions are as follows:

\begin{align}
  L_{rec} &= ||x-G(x)||_1 \\
  L_{adv} &= ||D(G(x))-1||_2^2 \\
  L_G &= \lambda_{rec}L_{rec} + \lambda_{adv}L_{adv} \\
  L_D &= ||D(x)-1||_2^2 + ||D(G(x))-0||_2^2
\end{align}

@@

\begin{align}
  \lambda_{adv}(n) = \begin{cases}
    0.001, & \text{if }n\leq10 \\
    0.005, & \text{if }10<n\leq30 \\
    0.015, & \text{otherwise}
  \end{cases}
\end{align}

@@

\begin{table}[h] 
  \centering  
    \begin{tabular}{lrrr}
    \hline
    \textbf{Model} & \textbf{L1} & \textbf{MSE} & \textbf{Adversarial}\\ 
    \hline \hline
      Baseline & $0.0956$ & $0.047$ & $0.1278$ \\
      Local Discriminator & $0.897$ & $1.044$ & $0.1014$ \\
      Residual Encoder & $0.08$ & $0.7814$ & $0.0941$ \\
    \hline
    \end{tabular}

  \caption{TODO} 
  \label{tab:4}
\end{table}

\begin{figure*}
	\captionsetup[subfigure]{labelformat=empty}
    \centering
    \begin{subfigure}[b]{0.175\textwidth}
        \includegraphics[width=\textwidth]{figs/fig2/masked_input}
        \caption{Masked input}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.175\textwidth}
        \includegraphics[width=\textwidth]{figs/fig2/baseline}
        \caption{Baseline}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.175\textwidth}
        \includegraphics[width=\textwidth]{figs/fig2/local}
        \caption{Discriminator}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.175\textwidth}
        \includegraphics[width=\textwidth]{figs/fig2/residual}
        \caption{Residual}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.175\textwidth}
        \includegraphics[width=\textwidth]{figs/fig2/ground_truth}
        \caption{GT}
    \end{subfigure}
  	\caption{Comparison of results for different model variants.}
  	\label{fig:fig2}
\end{figure*}

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figs/fig3/fig3}
        \caption{
        	\begin{tabular}{ccc}
        	Masked input & Local & GT
        	\end{tabular}
        }
    \end{subfigure}
  	\caption{Outpainting example using local discriminator model.}
  	\label{fig:fig3}
\end{figure}

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figs/fig4/fig4}
    \end{subfigure}
  	\caption{Outpainting result example using residual encoder model.}
  	\label{fig:fig4}
\end{figure}

\begin{figure}
	\captionsetup[subfigure]{labelformat=empty}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figs/fig5/fig5}
    \end{subfigure}
  	\caption{Example of anomaly with the hallucination.}
  	\label{fig:fig5}
\end{figure}
